---
title: "Chi-squared test on Lotto Max"
author: "Zhongmang Cheng"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# A First Look at the Histograms
Prior to proceeding with the test, we would like to have a general idea of shape of the statistics. 
A comparison has been created between generated winning numbers and observed winning numbers.

Generated winning numbers was created using a discrete uniform random variable with a expectation of 1/50.
Observed winning numbers was obtained from the official website of Lotto max: https://www.lottomaxnumbers.com/statistics

A first look at the data suggest that the observed data follows a normal distribution as expected, and that there is no abnormally.

```{r, echo = TRUE}
set.seed(2345)
random_numbers <- sample(1:50, size = 3080, replace = TRUE)

generated_frequency <- table(random_numbers)
observed_frequency <- c(61, 62, 56, 62, 67, 53, 61, 54, 71, 57, 61, 56, 55, 70, 60, 54, 62, 65, 73, 51, 62, 66, 56, 60, 63, 70, 58, 63, 58, 64, 66, 71, 54, 55, 62, 69, 67, 69, 71, 68, 53, 57, 59, 65, 57, 73, 51, 66, 56, 60)

par(mfrow = c(1, 2))

hist(generated_frequency, xlim = range(30, 90))
hist(observed_frequency, xlim = range(30, 90))
```


# Chi-squared Test for Fair Game
We propose the following null hypothesis and alternative hypothesis:

H0: the lottery game is fair, the numbers are drawn randomly and without bias.\
HA: the lottery game is not fair, there is some form of bias or manipulation in the drawing process.

Chi-squared Test will be used to test our hypothesis. 

```{r, echo = TRUE}
expected_frequency <- rep(1/50, 50)
observed_frequency <- c(61, 62, 56, 62, 67, 53, 61, 54, 71, 57, 61, 56, 55, 70, 60, 54, 62, 65, 73, 51, 62, 66, 56, 60, 63, 70, 58, 63, 58, 64, 66, 71, 54, 55, 62, 69, 67, 69, 71, 68, 53, 57, 59, 65, 57, 73, 51, 66, 56, 60)

chi_squared_test <- chisq.test(observed_frequency, p = expected_frequency)

print(chi_squared_test)
```

**The result of the test indicates a p-value greater than the critical value (0.95).**
**We failed to reject the null hypothesis (H0).**

This suggest that the numbers are very likely drawn randomly and without bias, and that Lotto Max is fair game.

